NGO analysis

### 2022-07-28 Dashboard/NGO Analysis dashboard alpha.Rmd
Adding further Excel testing.

```{r sidebar_button}
# download button
uiOutput("downloadUI")
```

@@

output$downloadUI <- renderUI({
  downloadButton("report", "Download report")
})

## Download and parameters -----------------------------------------------------
output$report <- downloadHandler(

@@

  # Data frames ================================================================

  # creating data frame from html file <- Can be deleted
  htmldata <- reactive({
    # validating file
    validate(need(
      tools::file_ext(input$html$datapath) == "html",
      "Please upload an html file."
    ))
    # Creating data frame
    read_proff(
      input$html$datapath,
      as.character(input$start), as.character(input$end)
    )
  })

  # Creating data frame from Excel file
  exceldata <- reactive({
    # validating file
    validate(need(
      tools::file_ext(input$excel$datapath) == "xlsx",
      "Please upload an Excel file."
    ))
    # Creating data frame
    widen_excel(input$excel$datapath)
  })

  # NGO title ==================================================================
  NGO <- reactive({
    # validating the file again
    validate(need(
      tools::file_ext(input$excel$datapath) == "xlsx",
      "Please upload an Excel file."
    ))
    # fetching the name
    read_excel(input$excel$datapath) |>
      # Turning the needed column into a list
      pull(Firm) |>
      # discarding all NAs in case the name is somewhere
      purrr::discard(is.na) |>
      # choosing the first available name
      first()
  })

@@

@@

  # Data frames ================================================================
  # Recreating, to make available to this chunk as well

  # creating data frame from html file
  htmldata <- reactive({
    # validating file
    validate(need(
      tools::file_ext(input$html$datapath) == "html",
      "Please upload an html file."
    ))
    # Creating data frame
    read_proff(
      input$html$datapath,
      as.character(input$start), as.character(input$end)
    )
  })

  # Creating data frame from Excel file
  exceldata <- reactive({
    # validating file
    validate(need(
      tools::file_ext(input$excel$datapath) == "xlsx",
      "Please upload an Excel file."
    ))
    # Creating data frame
    widen_excel(input$excel$datapath)
  })

@@

## Function to read Excel file -------------------------------------------------
widen_excel <- function(file) {
  read_excel(file) |>
    # uniting Category and Name to deal with multiple 'Other'
    unite(
      entry, c(Category, Name),
      sep = "_"
    ) |>
    # deleting Sub-category and Firm
    select(-`Sub-category`, -Firm) |>
    # pivoting to create 'tidy' variables
    pivot_wider(names_from = entry, values_from = Value) |>
    # cleaning names
    clean_names()
}

@@

# mutate(
#       valutakode = ifelse(year == 2019, "USD", valutakode),
#       leder_annen_godtgjorelse = NA,
#       lederlonn = NA,
#       arsresultat = NA,
#       sum_egenkapital_og_gjeld = NA
#     ) |>

### Test version of function that creates multiple data problems----------------
# read_proff <- function(file, start, end) {
#   read_html(file, encoding = "UTF-8") |>
#     html_node("#inner-frame") |>
#     html_table() |>
#     select( # removing empty column
#       REGNSKAPSPERIODE:all_of(start)
#     ) |> # using all_of()
#     mutate( # removing whitespace
#       REGNSKAPSPERIODE = str_squish(REGNSKAPSPERIODE)
#     ) |>
#     filter( # removing duplicate table - no idea why it's necessary
#       !row_number() > 176
#     ) |>
#     filter( # removing duplicate rows
#       !grepl("Lukk", REGNSKAPSPERIODE)
#     ) |>
#     pivot_longer( # tidying data
#       all_of(end):all_of(start),
#       names_to = "year"
#     ) |> # using all_of()
#     mutate( # changing to real NAs and turning years into numbers
#       value = na_if(value, "-"),
#       year = as.integer(year),
#       valutakode = ifelse(REGNSKAPSPERIODE == "Valutakode", value, NA)
#     ) |>
#     fill( # filling the currency column
#       valutakode,
#       .direction = "updown"
#     ) |>
#     filter( # removing dates and redundant currency
#       REGNSKAPSPERIODE != "Sluttdato" & REGNSKAPSPERIODE != "Startdato" &
#         REGNSKAPSPERIODE != "Valutakode"
#     ) |>
#     mutate( # removing whitespace in numbers
#       value = str_replace_all(value, "\\s", ""),
#       value = as.numeric(value),
#       value = ifelse(grepl("i hele 1000", REGNSKAPSPERIODE), NA, value)
#     ) |>
#     distinct() |> # removing any remaining duplicate rows
#     pivot_wider( # pivoting
#       names_from = REGNSKAPSPERIODE, values_from = value
#     ) |>
#     clean_names() |> # tidying names
#     select(
#       -lederlonn_i_hele_1000,
#       -resultatregnskap_i_hele_1000,
#       -balanseregnskap_i_hele_1000
#     ) |> # removing headings
#     rename( # making it clear what lonn refers to
#       lederlonn = lonn
#     ) |>
#     # Creating new variables for plots
#     mutate(
#       annual_profit = ifelse(arsresultat >= 0, TRUE, FALSE),
#       balance_positive = ifelse(sum_egenkapital_og_gjeld >= 0, TRUE, FALSE),
#     ) |>
#     # bad test - REMOVE
#     mutate(
#       valutakode = ifelse(year == 2019, "USD", valutakode),
#       leder_annen_godtgjorelse = NA,
#       lederlonn = NA,
#       arsresultat = NA,
#       sum_egenkapital_og_gjeld = NA
#     ) |>
#     arrange(desc(year)) # arranging by year
# }
```

>>>>

```{r sidebar_button}
# download button - using uiOutput() because of an apparent bug
uiOutput("downloadUI")
```

@@

## Download button -------------------------------------------------------------
output$downloadUI <- renderUI({
  downloadButton("report", "Download report")
})

## Report download and parameters ----------------------------------------------s
output$report <- downloadHandler(

@@

  # Excel data frame ===========================================================

  ## Validating and creating long version of Excel file, if possible------------
  excel_test <- reactive({
    # validating file
    validate(need(
      tools::file_ext(input$excel$datapath) == "xlsx",
      "Please upload an Excel file."
    ))

    widen_excel_test(input$excel$datapath)
  })

  ## Creating the primary Excel data frame--------------------------------------
  exceldata <- reactive({
    widen_excel(excel_test())
  })

  # # Creating data frame from Excel file
  # exceldata <- reactive({
  #   # validating file
  #   validate(need(
  #     tools::file_ext(input$excel$datapath) == "xlsx",
  #     "Please upload an Excel file."
  #   ))
  #   # Creating data frame
  #   widen_excel(input$excel$datapath)
  # })

  # NGO title ==================================================================
  NGO <- reactive({
    # checking for the required column again
    if ("firm" %in% names(excel_test())) {
      # fetching the name
      excel_test() |>
        # Turning the needed column into a list
        pull(firm) |>
        # discarding all NAs in case the name is somewhere
        purrr::discard(is.na) |>
        # choosing the first available name
        first()
    } else {
      validate("Please upload data using the specified Excel template.")
    }
  })

@@

  # NGO <- reactive({
  #   # validating the file again
  #   validate(need(
  #     tools::file_ext(input$excel$datapath) == "xlsx",
  #     "Please upload an Excel file."
  #   ))
  #   # fetching the name
  #   read_excel(input$excel$datapath) |>
  #     # Turning the needed column into a list
  #     pull(Firm) |>
  #     # discarding all NAs in case the name is somewhere
  #     purrr::discard(is.na) |>
  #     # choosing the first available name
  #     first()
  # })
  # 
  # title <- reactive({
  #   # testing if NGO() is NA
  #   if (is.na(NGO())) {
  #     # Printing warning if NA
  #     "Name of NGO missing"
  #     # Printing name of NGO
  #   } else {
  #     NGO()
  #   }
  # })

@@

  # Data frames ================================================================
  ## creating data frame from html file-----------------------------------------
  htmldata <- reactive({
    # validating file
    validate(need(
      tools::file_ext(input$html$datapath) == "html",
      "Please upload an html file."
    ))
    # Creating data frame
    read_proff(
      input$html$datapath,
      as.character(input$start), as.character(input$end)
    )
  })

  # Recreating, to make available to this chunk as well
  ## Validating and creating long version of Excel file, if possible------------
  excel_test <- reactive({
    # validating file
    validate(need(
      tools::file_ext(input$excel$datapath) == "xlsx",
      "Please upload an Excel file."
    ))

    widen_excel_test(input$excel$datapath)
  })

  ## Creating the primary Excel data frame--------------------------------------
  exceldata <- reactive({
    widen_excel(excel_test())
  })

@@

## Functions to read Excel file ------------------------------------------------
widen_excel_test <- function(file) {
  # Reading Excel file
  read_excel(file) |> 
    # cleaning names to enable later operations
    clean_names()
}

# Function to create primary Excel data frame
widen_excel <- function(object) {
  # testing if all columns are present
  if (setequal(
    names(object),
    c("year", "firm", "category", "sub_category", "name", "value")
  )) {
    object |>
      # uniting category and name to deal with multiple 'Other'
      unite(
        entry, c(category, name)
      ) |>
      # deleting Sub-category and Firm
      select(-sub_category, -firm) |>
      # pivoting to create 'tidy' variables
      pivot_wider(names_from = entry, values_from = value) |>
      # cleaning names
      clean_names()
    # Notfification if all columns are not present
  } else {
    validate("Please upload data using the specified Excel template.")
  }
}

# widen_excel <- function(file) {
#   read_excel(file) |>
#     # uniting Category and Name to deal with multiple 'Other'
#     unite(
#       entry, c(Category, Name),
#       sep = "_"
#     ) |>
#     # deleting Sub-category and Firm
#     select(-`Sub-category`, -Firm) |>
#     # pivoting to create 'tidy' variables
#     pivot_wider(names_from = entry, values_from = Value) |>
#     # cleaning names
#     clean_names()
# }

@@

# mutate(
#       valutakode = ifelse(year == 2019, "USD", valutakode),
#       leder_annen_godtgjorelse = NA,
#       lederlonn = NA,
#       arsresultat = NA,
#       sum_egenkapital_og_gjeld = NA
#     ) |>
```
